---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

On January 28, 1986, a routine launch was anticipated for the Challenger space shuttle. Seventy-three
seconds into the flight, disaster happened: the shuttle broke apart, killing all seven crew members on
board. An investigation into the cause of the disaster focused on a critical seal called an O-ring, and
it is believed that damage to O-rings during shuttle launches may have been related to the ambient
temperature at the time of the launch. Using data from 24 shuttle missions, we fit a simple logistic
regression model. Here, Temperature is the temperature in Fahrenheit at launch time and the response
variable, Failure, is an indicator of O-rings failures (Failure = 1 if there was at least O-ring that failed.
Here is a summary of the logistic regression model fit to these data:


            Estimate Std. Error z value Pr(> |z|)
(Intercept)   10.88     5.70     1.91     0.06
Temperature   -0.17     0.08    -2.05     0.04


1. Find a 90% confidence interval and a 95% confidence interval for the coefficient on Temperature. Which one of the intervals is wider? Can you provide some intuition on why?

```{r}

estimate <- -0.17
zval_90 <- qnorm(0.95) # 90% two-tailed confidence interval
zval_95 <- qnorm(0.975)
stdError <- 0.08

confLower_90 <- estimate - zval_90*0.08
confUpper_90 <- estimate + zval_90*0.08

confLower_95 <- estimate - zval_95*0.08
confUpper_95 <- estimate + zval_95*0.08

print(paste('90%:', confLower_90, confUpper_90))

print(paste('95%:', confLower_95, confUpper_95))

```

## The 95% confidence interval is wider because we are casting a larger statistical net. We are saying "We are 95% confident the paramenter of interest is in this range" VS 90% confident. As we become more confident, we increase the range.


2. In the context of the problem, interpret the 90% confidence interval you found in question 1, on the
log-odds scale.


```{r}

logodds_lower <- exp(confLower_90)
logodds_Upper <- exp(confUpper_90)
print(paste('log odds',logodds_lower,logodds_Upper ))

```
## Here we would still reject the null hypothesis that the estimate of temperature is different than zero.

3. Find a 90% confidence interval for the effect of a one-degree increase in temperature on the odds of O-ring failure and interpret your result.

```{r}

#Exp of the estimate value to find the odds odds
change <- exp(-0.17)

conflower_90 <- change - zval_90*0.08
confupper_90 <- change + zval_90*0.08

print(paste('90%:', conflower_90, confupper_90))

```
## We are 90% confident that the effect of a one-degree increase of temperature on the odds of o-ring failure will be between 71.2% and 97.5%


4. Find the probability of O-ring failure when the ambient temperature is 52◦F and when the ambient
temperature is 72◦F. Which temperature seems more favorable for a successful launch?


```{r}

probFail_52 = exp(10.88 -0.17*52)

probFail_72 = exp(10.88 -0.17*72)

print(paste('failure probability for 52 and 72 degrees F:', probFail_52, probFail_72))

```
## The probability of failure at 52 degrees F is over 1, therefore is much less favorable than the 26% of 72 degrees F.

5. Can you find the coldest temperature so that the probability of O-ring failure is less than 10%?



```{r}
for (i in 60:100){
  
  val = exp(10.88 -0.17*i)
  print(paste('Temp, Percent:', i, round(val,3)))
}
#print(paste('failure probability for 52 and 72 degrees F:', probFail_52, probFail_72))

```

## The first temperature with less than 10% failure rate is 78 degrees F